---
title: "Intro to R - Week 1"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to R for Data Science (FutureLearn) - Week 1

By Graeme West

This notebook covers the material in Week 1 of [Introduction to R for Data Science](https://www.futurelearn.com/courses/data-science/5) by Purdue University on FutureLearn.

The subject matter of the data is airline flight data from the United States. Week 1 covers basic setup, exploratory data analysis, selecting data, and basic plotting.

## Step 1.12

Read in the basic 2008 data (this takes a couple of minutes):
```{r}
myDF <- read.csv("./datasets/2008.csv.bz2")
```


Phew, that was tough! Let's try and make heads and tails of it...
```{r}
head(myDF)
```

```{r}
tail(myDF)
```

## Step 1.13: Looking at only Indianapolis flights

First let's look at the top of the 'Origin' column.

```{r}
head(myDF$Origin)
```
And the tail:
```{r}
tail(myDF$Origin)
```

Let's also check the 'dest' column:
```{r}
head(myDF$Dest)
```

```{r}
tail(myDF$Dest)
```


## Step 1.15: Identifying Properties

Let's slice by origin to identify flights that began in Indianapolis:

```{r}
head(myDF$Origin == "IND")
```
Okay, so that returned a series of Booleans. Let's get the total number by doing a 'sum':

```{r}
sum(myDF$Origin == "IND")
```
Let's count the flights that were destined to Indianapolis:
```{r}
sum(myDF$Dest == "IND")
```

## Step 1.17: Extracting Flight Data with a Common City of Origin

Make a new DataFrame with the entries from the total data where the origin airport was Indianapolis:
```{r}
MyIndyOrigins <- subset(myDF, myDF$Origin == "IND")
head(MyIndyOrigins)
```

Let's do the same for the flights that were destined for Indianapolis:
```{r}
IndianapolisDestinations <- subset(myDF, myDF$Dest == "IND")
head(IndianapolisDestinations)
```

How would we determine which flights departed in a particular month? We'll try a table first.

```{r}
table(MyIndyOrigins$Month)
```

And now a table:
```{r}
plot(table(MyIndyOrigins$Month))
```

Now the same for destinations:
```{r}
plot(table(IndianapolisDestinations$Month))
```

## Step 1.19: Analyzing the Departure Times of Flights

Let's look at the flights with departure times before 6AM:

```{r}
sum(MyIndyOrigins$DepTime < 600)
```
This causes an issue because some value are NaNs. We'll set the `na.rm` parameter to drop these from the calculation:
```{r}
sum(MyIndyOrigins$DepTime < 600, na.rm=TRUE)
```

Let's count the 'NaN' values in the `DepTime` column:
```{r}
sum(is.na(MyIndyOrigins$DepTime))
```

## Step 1.21 Annotating R Code with Comments

"Make a plot of departure times, grouped by hour, for the flights from ATL to LAX.":

```{r}
atllaxflights <- myDF[ which(myDF$Origin == "ATL" & myDF$Dest == "LAX"), ]
head(atllaxflights)
```
I noticed that there are several NA entries in the DepTime column for this DataFrame. So let's drop those, then group it by hour:
```{r}
atllaxflights_clean <- na.omit(atllaxflights, cols="DepTime")
atllaxflights_clean$timestamp <- strptime(sprintf("%04d", atllaxflights_clean$DepTime), format="%H%M")
atllaxflights_clean$depHour <- as.numeric(format(atllaxflights_clean$timestamp, format="%H"))
head(atllaxflights_clean, 250)
```

Now, let's generate a plot of these, grouped by hour:
```{r}
hist(atllaxflights_clean$depHour, main="ATL to LAX Flights by Hour of Day, 2008", xlab="Hour of Day", ylab="Number of Flights", border="blue")
```

