---
title: "Introduction to R for Data Science (FutureLearn)"
author: "Graeme West"
date: "13/03/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R for Data Science (FutureLearn)

By Graeme West

This notebook covers the material in Week 1 of [Introduction to R for Data Science](https://www.futurelearn.com/courses/data-science/5) by Purdue University on FutureLearn.

This repository is intended to serve as a record of my learning, and my solutions to the various challenges set by the instructors.

The subject matter of the data is airline flight data from the United States.

The original dataset for week 1 is too large to store on GitHub, so here's a handy R snippet that will download it for you:
```{r, warning=FALSE}
download.file("http://stat-computing.org/dataexpo/2009/2008.csv.bz2",destfile="datasets/2008.csv.bz2",method="libcurl")
```


# Week Index

1. [Week 1 (Introduction, setup, EDA, basic plotting)](Intro_to_R_Week_1.md)
2. [Week 2 ](Intro_to_R_Week_2.md)

**FURTHER WEEKS - TBD**